Early Observations of the Type Ia Supernova \abc:\\ A Case of Interaction with Nearby, Unbound Material and/or Strong Ejecta Mixing

### 2018-05-22 doc/ms.tex
final corrections from proofs

We have presented observations of the extraordinarily early discovery of the 
normal Type Ia supernova \abc. Our fast-response follow-up 
campaign allowed us to draw the following conclusions:

\begin{itemize}

    \item Extrapolation of the early light curve shows that the initial
    detection of \abc\ occurred only $0.15\pm_{0.07}^{0.15}$~d after the time
    of first light, $t_0$.

@@

    \item The strong and short-lived carbon features seen in the earliest
    spectra of \abc\ can only be explained if there is incomplete burning. The
    pulsational delayed detonation models presented in
    \citet{2014MNRAS.441..532D} produce \ion{C}{2} absorption that is as
    strong as \ion{Si}{2} at very early phases.

    \item In contrast to SN\,2011fe, $(B - V)_0$ is $\sim$0.5 mag bluer for
    \abc\ at $t \approx -16$\,d. Furthermore, the $(B - V)_0$ colors of \abc\
    show no evolution over the first $\sim$7\,d of observations.

    \item Finally, we show that the early light curve evolution and colors of
    \abc\ are best matched by the pulsational delayed detonation
    models of \citet{2014MNRAS.441..532D} and the ejecta-mixing models of
    \citet{2016ApJ...826...96P}.

\end{itemize}
%
Taken together, these observations suggest that the early emission
from \abc\ is due to either the collision of the SN with nearby, unbound
material and/or that there is significant mixing of $^{56}$Ni in the SN
ejecta. The PDD models from \citet{2014MNRAS.441..532D} are particularly
attractive for explaining \abc, because they produce strong
\ion{C}{2} absorption at early times. In the future, it would be useful to
investigate more detailed PDD models that incorporate strong $^{56}$Ni mixing
to see if they better replicate the observations of \abc\ as it is otherwise
difficult to distinguish between these two scenarios.

Extremely early observations of young SNe provide a ``smoking
gun'' to probe the mixing level in the ejecta, which, in turn, is 
a result of the explosion mechanism. Wide-field, high-cadence surveys, such as the Zwicky Transient Facility \citep{2016PASP..128h4501B} and ATLAS \citep{2011PASP..123...58T,2013RSPTA.37120269T}, will discover a large number of very young SNe over the next few years, allowing us to extend our studies beyond single objects. While the sample of extremely young SNe Ia will grow by more than an order of magnitude, the detection of shock breakout cooling and ejecta--companion interaction will prove challenging. Given the diminutive size of WDs, the thermal emission following shock breakout can only be detected to $\sim 10\,\mathrm{Mpc}$ on 1-m class telescopes. Furthermore, only $\sim$10\% of single-degenerate progenitors are expected to give rise to detectable emission following the collision of the SN ejecta with the binary companion \citep{2010ApJ...708.1025K}. Despite these limitations, this study of \abc\ shows that the early detection of SNe Ia may probe explosion physics by measuring the amount of mixing in the SN ejecta. Moving forward, a large sample of such objects will enable strict constraints on the proposed explosion mechanisms for Type Ia SNe.

Finally, we close by emphasizing the importance of fast-response photometric and spectroscopic follow-up campaigns. Without the early recognition of the youth of this SN, and the associated follow-up, much of the analysis presented herein would not have been possible. The ability to trigger such observations is essential to improve our physical understanding of Type Ia SNe.

@@

AAM is funded by the Large Synoptic Survey Telescope Corporation in support of
the Data Science Fellowship Program. YC thanks support from a postdocal
fellowship at the eScience Institute, University of Washington.

@@

The Intermediate Palomar Transient Factory project is a scientific collaboration among the California Institute of Technology, Los Alamos National Laboratory, the University of Wisconsin, Milwaukee, the Oskar Klein Center, the Weizmann Institute of Science, the TANGO Program of the University System of Taiwan, and the Kavli Institute for the Physics and Mathematics of the Universe. This work was supported by the GROWTH project funded by the National Science Foundation under Grant No 1545949. Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the NASA. This work makes use of observations from the LCO network. These results made use of the Discovery Channel Telescope at Lowell Observatory. Lowell is a private, non-profit institution dedicated to astrophysical research and public appreciation of astronomy and operates the DCT in partnership with Boston University, the University of Maryland, the University of Toledo, Northern Arizona University and Yale University. The upgrade of the DeVeny optical spectrograph has been funded by a generous grant from John and Ginger Giovale. Based on observations made with the Nordic Optical Telescope, operated by the Nordic Optical Telescope Scientific Association at the Observatorio del Roque de los Muchachos, La Palma, Spain, of the Instituto de Astrof\'isica de Canarias.

\facility{DCT, Gemini:Gillett, Hale, Keck:I, Keck:II, LCOGT, PO:1.2\,m, PO:1.5\,m, NOT, VLT, Swift, OANSPM:HJT}

>>>>

We have presented observations of the extraordinarily early discovery of the 
normal SN Ia \abc. Our fast-response follow-up 
campaign allowed us to draw the following conclusions:

\begin{enumerate}

    \item Extrapolation of the early light curve shows that the initial
    detection of \abc\ occurred only $0.15\pm_{0.07}^{0.15}$~days after the
    time of first light, $t_0$.

@@

    \item The strong and short-lived carbon features seen in the earliest
    spectra of \abc\ can only be explained if there is incomplete burning.
    The pulsational delayed-detonation models presented in
    \citet{2014MNRAS.441..532D} produce \ion{C}{2} absorption that is as
    strong as \ion{Si}{2} at very early phases.

    \item In contrast to SN\,2011fe, $(B - V)_0$ is $\sim$0.5\,mag bluer for
    \abc\ at $t \approx -16$\,days. Furthermore, the $(B - V)_0$ colors of
    \abc\ show no evolution over the first $\sim$7\,days of observations.

    \item Finally, we show that the early light-curve evolution and colors
    of \abc\ are best matched by the pulsational delayed-detonation models
    of \citet{2014MNRAS.441..532D} and the ejecta-mixing models of
    \citet{2016ApJ...826...96P}.

\end{enumerate}
%
Taken together, these observations suggest that the early emission from
\abc\ is due to either the collision of the SN with nearby, unbound material
and/or that there is significant mixing of $^{56}$Ni in the SN ejecta. The
PDD models from \citet{2014MNRAS.441..532D} are particularly attractive for
explaining \abc, because they produce strong \ion{C}{2} absorption at early
times. In the future, it would be useful to investigate more detailed PDD
models that incorporate strong $^{56}$Ni mixing to see whether they better
replicate the observations of \abc, as it is otherwise difficult to
distinguish between these two scenarios.

Extremely early observations of young SNe provide a ``smoking gun'' to probe
the mixing level in the ejecta, which, in turn, is a result of the explosion
mechanism. Wide-field, high-cadence surveys, such as the Zwicky Transient
Facility \citep{2016PASP..128h4501B} and ATLAS
\citep{2011PASP..123...58T,2013RSPTA.37120269T}, will discover a large
number of very young SNe over the next few years, allowing us to extend our
studies beyond single objects. While the sample of extremely young SNe Ia
will grow by more than an order of magnitude, the detection of shock
breakout cooling and ejecta--companion interaction will prove challenging.
Given the diminutive size of WDs, the thermal emission following shock
breakout can only be detected to $\sim 10\,\mathrm{Mpc}$ on 1\,m class
telescopes. Furthermore, only $\sim$10\% of single-degenerate progenitors
are expected to give rise to detectable emission following the collision of
the SN ejecta with the binary companion \citep{2010ApJ...708.1025K}. Despite
these limitations, this study of \abc\ shows that the early detection of SNe
Ia may probe explosion physics by measuring the amount of mixing in the SN
ejecta. Moving forward, a large sample of such objects will enable strict
constraints on the proposed explosion mechanisms for SNe Ia.

Finally, we close by emphasizing the importance of fast-response photometric
and spectroscopic follow-up campaigns. Without the early recognition of the
youth of this SN and the associated follow-up, much of the analysis
presented herein would not have been possible. The ability to trigger such
observations is essential to improve our physical understanding of SNe Ia.

@@

AAM is funded by the Large Synoptic Survey Telescope Corporation in support
of the Data Science Fellowship Program. YC acknowledges support from a
postdoctoral fellowship at the eScience Institute, University of Washington.

@@

The Intermediate Palomar Transient Factory project is a scientific collaboration among the California Institute of Technology, Los Alamos National Laboratory, the University of Wisconsin, Milwaukee, the Oskar Klein Center, the Weizmann Institute of Science, the TANGO Program of the University System of Taiwan, and the Kavli Institute for the Physics and Mathematics of the Universe. This work was supported by the GROWTH project funded by the National Science Foundation under Grant No 1545949. Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the NASA. This work makes use of observations from the LCO network. These results made use of the Discovery Channel Telescope at Lowell Observatory. Lowell is a private, nonprofit institution dedicated to astrophysical research and public appreciation of astronomy and operates the DCT in partnership with Boston University, the University of Maryland, the University of Toledo, Northern Arizona University, and Yale University. The upgrade of the DeVeny optical spectrograph has been funded by a generous grant from John and Ginger Giovale. Based on observations made with the Nordic Optical Telescope, operated by the Nordic Optical Telescope Scientific Association at the Observatorio del Roque de los Muchachos, La Palma, Spain, of the Instituto de Astrof\'isica de Canarias.

\facility{DCT, Gemini:Gillett, Hale, Keck:I, Keck:II, LCOGT, PO:1.2\,m, PO:1.5\,m, NOT, VLT, \textit{Swift}, OANSPM:HJT}
